---
title: "3D plots"
author: "Bryce Pulver"
date: "September 26, 2019"
output: html_document
---

# Analyzing topographic change in the USA

```{r setup, include=FALSE}
library(tidyverse)
library(sf)  #Spatial datasets
library(mapview) #interactive mapping
library(raster) #Raster analysis
library(elevatr) #National Elevation Dataset downloader. 
library(rayshader) #3D plotting package
knitr::opts_chunk$set(echo = TRUE)
```


## Check out the data

```{r}
topo_change <- st_read('data/topo_change/topo_change_polygons.shp')
#Checkout the column names
#names(topo_change)
biggest_area <- topo_change %>%
  arrange(desc(AREA_SQ_KM)) %>%
  slice(1:10)
# Map all mines
mapview(biggest_area)
```

### Checkout mine in Arizona

```{r}
# Subset to just Arizona
az_mine <- biggest_area %>%
  filter(QUADNAME == 'esperanza_mill_AZ')
#Check that it is the right site
mapview(az_mine)
```


## Getting our own elevation datasets

### Download pre-mining DEM 

```{r}
#Check projection of az_mine
#st_crs(az_mine)
az_raster_before <- get_elev_raster(az_mine,z=12)
#Look at the structure of the data
#str(az_raster)
#Summary of the data
#summary(az_raster)
plot(az_raster_before)
#mapview(az_raster_before)
```



### Shrink the resolution

```{r}
small_az <- aggregate(az_raster_before,fact=6)
mapview(small_az)
```


### Take the slope of the data

```{r}
az_slope <- terrain(small_az,opt='slope',unit='degrees')
mapview(az_slope)
```


### Take the aspect of the data

```{r}
az_aspect <- terrain(small_az,opt='aspect',unit='degrees')
mapview(az_aspect)
```


## 3D plotting mechanism for elevation data

```{r}
az_matrix <- matrix(raster::extract(small_az,
                    raster::extent(small_az),buffer=500),
                    nrow=ncol(small_az), ncol=nrow(small_az))
az_matrix %>%
  sphere_shade(texture = 'desert') %>%
  add_water(detect_water(az_matrix),color='desert') %>%
  add_shadow(ray_shade(az_matrix,zscale=2),0.5) %>%
  add_shadow(ambient_shade(az_matrix)) %>%
  plot_3d(az_matrix,zscale=10,fov = 0,theta=135,zoom=0.75)
```


# In Class work


For this section we are really going to focus on the mine area
so first I'm going to buffer our az_mine polygon by 2km and then
clip both the before and after mining rasters to this new shape

```{r}
#Read in raster data
az_raster_after <- raster('data/srtm_14_06.tif')
az_mine_2km <- az_mine %>%
  st_transform(2163) %>%
  st_buffer(2000) %>%
  #Reproject into SRTM datum (WGS84)
  st_transform(st_crs(az_raster_after))
#Get extent of buffered mine area
aoi <- extent(az_mine_2km)
#Crop after to area of interest
mine_after <- crop(az_raster_after,aoi)
#Crop before to area of interest (need to also reproject)
mine_before <- crop(az_raster_before %>%
                      projectRaster(.,crs=projection(mine_after)),
                    aoi) %>%
  projectRaster(.,mine_after) #match resolution
#Force rasters to be exactly same size
mine_before_same <- trim(mine_before)
mine_after_same <- crop(mine_after,mine_before_same)
#Stack these things in a rasterBrick (which we can do because we 
#made them the exact same size)
mines <- brick(mine_before_same,mine_after_same)
#Add a dem_difference layer
mines[[3]] <- mines[[1]] - mines[[2]]
#Rename raster layers
names(mines) <- c('Pre_mining','Post_mining','DoD')
plot(mines)
```


All data for this section should be the data clipped to the mining area 
(generated in the code chunk above)

## 1) Get the slope of the SRTM (post-mining) dataset and plot this data (use the data that is clipped to the mine only).


```{r}
az_slope_post <- terrain(mines$Post_mining,opt='slope',unit='degrees')
mapview(az_slope_post)
```


## 2) Use the brick code from above to plot slopes for the pre-mining, post-mining, and difference of slope

```{r}
az_slope_pre <- terrain(mines$Pre_mining,opt='slope',unit='degrees')
az_slope_dif <- terrain(mines$DoD,opt = 'slope',unit = 'degrees')
mines_slope <- brick(az_slope_pre,az_slope_post,az_slope_dif)
names(mines_slope) <- c('Pre_mining_slope','Post_mining_slope','DoD_slope')
plot(mines_slope)
```


## 3) Use rayshader to make a 3d plot of the pre-mining mine

```{r}
az_matrix_pre <- matrix(raster::extract(mines$Pre_mining,
                    raster::extent(mines$Pre_mining),buffer=500),
                    nrow=ncol(mines$Pre_mining), 
                    ncol=nrow(mines$Pre_mining))
az_matrix_pre %>%
  sphere_shade(texture = 'desert') %>%
  add_water(detect_water(az_matrix_pre),color='desert') %>%
  add_shadow(ray_shade(az_matrix_pre,zscale=2),0.5) %>%
  add_shadow(ambient_shade(az_matrix_pre)) %>%
  plot_3d(az_matrix_pre,zscale=10,fov = 0,theta=135,zoom=0.75)
```


## 4) Use rayshader to make 3d plot of the post-mining mine

```{r}
az_matrix_post <- matrix(raster::extract(mines$Post_mining,
                    raster::extent(mines$Post_mining),buffer=500),
                    nrow=ncol(mines$Post_mining), 
                    ncol=nrow(mines$Post_mining))
az_matrix_post %>%
  sphere_shade(texture = 'desert') %>%
  add_water(detect_water(az_matrix_post),color='desert') %>%
  add_shadow(ray_shade(az_matrix_post,zscale=2),0.5) %>%
  add_shadow(ambient_shade(az_matrix_post)) %>%
  plot_3d(az_matrix_post,zscale=10,fov = 0,theta=135,zoom=0.75)
```

Plotting the difference of slope in a 3d plot

```{r}
az_matrix_dif <- matrix(raster::extract(mines_slope$DoD_slope,
                    raster::extent(mines_slope$DoD_slope),buffer=500),
                    nrow=ncol(mines_slope$DoD_slope), 
                    ncol=nrow(mines_slope$DoD_slope))
az_matrix_dif %>%
  sphere_shade(texture = 'desert') %>%
  add_water(detect_water(az_matrix_dif),color='desert') %>%
  add_shadow(ray_shade(az_matrix_dif,zscale=2),0.5) %>%
  add_shadow(ambient_shade(az_matrix_dif)) %>%
  plot_3d(az_matrix_dif,zscale=10,fov = 0,theta=135,zoom=0.75)

```

